### 2018-08-14 review_response.tex
minor comments from 1st reviewer

\documentclass[]{scrartcl}
\usepackage{csquotes}
\MakeOuterQuote{"}

%opening
\title{Response for SCSS-JSC submission 2}
\subtitle{Aligning Concepts across Proof Assistant
Libraries}
\author{Thibault Gauthier \and Cezary Kaliszyk}

\usepackage{blindtext} 
\usepackage{framed}
%\usepackage[parfill]{parskip}

\setlength{\parskip}{1.5mm}
\setlength{\parindent}{0mm}

\begin{document}

\maketitle

We thank the reviewers for the valuable comments and questions. We have 
attempted to correct all
the issues suggested in the reviews and we give further answers to the 
questions raised in this response letter.

\section*{Reviewer 1}

\begin{leftbar}
My main concern about the paper is related to the usability by people that are 
not the authors of the paper. In several places of the paper, the authors 
indicate that some steps of the alignment process are carried out manually, but 
this is not completely clear; so, the authors should clarify this point in 
order to know which steps are conducted manually and which steps are automated. 
Moreover, the output results are huge plain text file, and exploring them is 
really time-consuming (as indicated previously, the authors have already done 
this, but it might be difficult to the same by other users in a different 
context). In addition, the output results must be manually analysed by the user.

Related to the previous point, the authors provide the final data obtained in 
their
experiments; however, they do not provide the programs that they have used. 
Hence, it is not possible to reproduce the authors' results or use their 
algorithms in different libraries and contexts.
\end{leftbar}

We have now provided the source code of all the experiments performed together 
with instructions how to rerun them
to ease the repeatability and reusability of our project.

When it comes to the section on human advice, the purpose is to allow the user 
to intervene in the decision process,
which could allow the algorithms to continue with more precise matches. Indeed, 
such a hybrid method is very
time consuming and was not evaluated in its full extent.
Except for the logical mappings, all presented matches were automatically 
derived.
We have now explained this in more detail in Section 5.4.

\begin{leftbar}Another issue that arose me when reading the paper were some of 
the applications of this work: translating results and conjecturing new 
theorems. Since the alignment of concepts is based on the quantity (and 
quality) of shared properties, how many results do you need to prove in order 
to discover the alignment of concepts? This is probably one drawback of your 
work: the need of big libraries (it is probably not a problem using the already 
developed libraries, but it might be a problem when creating a new library).
\end{leftbar}

Indeed, the proposed approach would not work well with small proof libraries.
However, given the bridges we constructed between all the considered libraries, 
a relatively small development may
map to specific structures in other libraries and allow the import of related 
theorems.

\begin{leftbar}Another point is whether you have noticed that some matchings 
are missed.
I mean, two concepts that were intended to represent the same mathematical 
object, but that are not matched. I guess that you do not have an automatic way 
of discovering those missing matches.
\end{leftbar}

The only proposed method of discovering such missing matches that our algorithm 
supports is through transitive matches.
We have now added some comments about it in section 4.6.

\begin{leftbar}It is a bit strange that you have not included the alignments in 
the same ITP, have you tried that experiment? Have you discovered something 
interesting? As an idea for the future, you could create some kind of alert 
system that inform the user about the matching of concepts in the libraries of 
the ITP to avoid duplications.
\end{leftbar}

We have tried experiments within a single ITP library. For curated libraries 
where no exact duplicates
occur (like the 6 libraries we consider) the alignments found are often 
concepts that are dual. These
include union and intersection of sets in Mizar, or addition and substraction 
in HOL libraries or
closely related concepts like lipschitzian and continuous.

In a work related to your question (presented at CICM Work in Progress last 
year), we tried to find the most likely substitutions (which mappings
should be applied together) and create a large number of conjectured formulas. 
Only 5 percent of them can be reproved. However they have not been too useful 
for current automation methods. This is now also shortly discussed in the paper.

\begin{leftbar}I wonder if you could define some threshold to differentiate the 
optimal matches, approximate matches and singular matches.
\end{leftbar}

Our current method is statistical, therefore the thresholds are estimates 
learned from previous library alignments.
We could also try to prove equivalence (modulo other mappings) of two similar 
concept but this would most of the times fail even for optimal matches because 
of minor differences in the definitions.

\begin{leftbar}
From my point of view, a graph presenting the workflow that is followed in your 
work to
align concepts might be useful for the reader.
\end{leftbar}
We included a graph together with clarifications to the overview of the 
procedure in Section 1.5.

\section*{Reviewer 2}

\subsection*{COMMENTS ABOUT THE RELEVANCE OF THE PROBLEM}

\begin{leftbar}
Automatically generating alignments is an important practical precondition to 
tackle the harder task of transferring knowledge between mathematical libraries 
at the shallow level, i.e. by translating a constant to the corresponding 
constant in the other library (avoiding to obtain two related but not 
identified notions). The problem seems therefore of primary importance.

However, the authors tackle the problem of aligning existing libraries from 
scratch. This needs to be done only once and forall and the automatically found 
result can be throughly corrected and improved by humans. After it is done, the 
most significant scenario, becomes the one where an author is starting a new 
formalization and he needs to pull in results from other libraries on the same 
constants he is interested in. This scenario is rather different for a series 
of reasons: 1) at the beginning of the new development there are usually only a 
small number of theorems proved and this makes it harder to find correct 
alignments; 2) the algorithm to automatically find alignments could be 
restricted to a subset of the system libraries that are already known to be 
potentially interesting.

I would like to hear from the authors what are their expectations on the 
behaviour of their algorithm on the new scenario. Maybe they could also run 
some benchmarks to receive a first impression, e.g. by showing how the matches 
improve when re-running the algorithm after adding one by one the lemmas in one 
theory.
\end{leftbar}

The proposed scenario is indeed quite different, and like the reviewer 
suggests, it would not work very well because of the first reason.
The only way to cope with it would be to automatically explore the theory 
before the alignment algorithm would be strong enough to discover new theorems 
by itself. As for the second reason why the proposed scenario is different, the 
restriction could be automatically found by a first pass of the alignment
algorithm.

Unfortunately for a number of the considered libraries no theory exploration 
mechanisms exist so far. Developing such mechanisms would in particular require 
powerful general purpose automation, which also has been limited in some of the 
systems. We completely agree that combining the presented work with theory 
exploration would be a very interesting line of work, but it would also be a 
large endeavor and has not been considered so far.

\subsection*{COMMENTS ABOUT THE TECHNICAL SOLUTION}
\begin{leftbar}All the choices the authors took seem reasonable. Of course, it 
is possible to think about totally different techniques to automatically 
identify alignments, for example borrowing other tools from AI and machine 
learning. However, being the first to address the problem at this scale, I 
believe they found a great compromise between simpliciy and efficiency. I will 
add later specific minor comments.

However, one remark I want to do now is that the kind of alignments the authors 
want to find should be directly related to the normalization they apply. For 
example, imagine that the authors want to align binary real division from HOL 
to ternary real division in some Coq libraries. Normalization should somehow 
``normalize away'' the proof argument in order to hope to match properties up 
to alpha-conversion. I can easily come up with other similar examples where 
aligning the library of a constructive, predicative type theory with that of an 
higher order logic requires to be more aggressive in the generalization process 
that produces properties. In previsous papers co-authored by the same authors 
certain classes of alignments were identified. I would have liked to see an 
explanation of normalization that is linked directly to those kind of 
alignments.
\end{leftbar}

Indeed, the manually constructed logical mappings are a weakness of our 
procedure.
However we have been able to find the logical mappings rather quickly, and they 
do not need to be extended as proof libraries grow.
It is true that for an algorithm to discover the logical mappings, the actual 
mappings would need to be more flexible.
In this respect, a possible extension of the algorithm could be to allow 
different parts of the term tree to be conceptualized.
However, we expect that the additional flexibility would reduce precision. We 
have included some discussion of this when considering
subterm conceptualizations.

\subsection*{COMMENTS ABOUT THE EVALUATION METHODOLOGY}

\begin{leftbar}
The evaluation of the technique is probably the weakest part of the paper and 
the one that suggests further future improvements. In particular, the algorithm 
only uses positive pre-existing knowledge in order to identify the alignments 
and complex heuristics to try to single out promising alignments and 
dependencies between the alignments. Finally, the responsibility of checking 
the quality of the alignments is totally left to the user.

A significant improvement would be to try to use automatic techniques to verify 
the quality of the alignments based on the generation of new knowledge, both 
positive and negative. For example, suppose that the system detect a possible 
alignment between concept A and B because A and B share a certain number of 
properties with a good score. In order to evaluate the alignment, the system 
could take for example all the properties on A that are not present for B and 
try to conjecture them to hold for B as well. Once the conjecture is done, 
automatic proof checkers and counterexample generators could be used to prove 
or disprove the conjecture. Even if they fail, the conjectures could often 
still be tested on a certain number of randomly generated tests. For example, 
it should not be hard to automatically understand that the alignment between 
natural and integer numbers is not optimal.
\end{leftbar}

This is a very good idea. There are three reasons why we have not been able to 
do this.
First, we do not have powerful enough general-purpose automation for all the 
provers considered.
Second, it is difficult to evaluate a match on its own. It makes little sense 
to map addition to multiplication but not 0 and 1 at the same time. Therefore 
we would need to consider and evaluate structural alignments consisting of a 
coherent set of matches.
Finally, given the number of matches evaluating each match using many  
automated reasoning experiments would be would be prohibitively time-consuming.

The closest to what the reviewer suggests and is currently possible has been 
considered by us in the CICM Work in Progress paper:
focusing on certain structural alignments one can evaluate the newly generated 
conjectures in some of the provers. The results
have been however much weaker than with other applications presented here. This 
is discussed in Section 1.5 and Section 5.

\begin{leftbar}
Coming to non optimal alignments, I would like to understand if the score 
computed by the algorithms is informative or not on the usefulness of the 
alignment. For example, aligning natural numbers with lists is quite useful 
because a natural number is a degenerate case of list (a list of elements of 
the unit type) and every theorem over lists induces a (possibly trivialized) 
theorem on natural numbers. Another example is aligning natural numbers with 
integers: even if the alignment is non optimal, the alignment becomes optimal 
(or almost so) if restricted to subsets of the two types (natural vs positive 
integers). Most other alignments, instead, are really likely to be completely 
useless.
\end{leftbar}

The scores are representative of the number and quality of the property shared 
by the two constants. So for a particular
pair of provers, the actual value of the score is indeed informative about the 
quality of the alignment, and the best
evidence of this are the actual correct alignments being found despite 
thousands of concepts present.

\begin{leftbar}
Did you ask a mathematician expert in the field what he thinks about the 
conjecture?
\end{leftbar}
We did not ask any expert mathematicians in the domain, instead relying of our 
own limited knowledge of the field to make those conjectures about the 
dynamical system.
We tried to prove those conjectures but failed a bit short so we did some 
empirical evaluation at dimension 2 confirming our intuition that the fix point 
we reach is the only stable fix point.

\begin{leftbar}
Minor comments
\end{leftbar}
We added examples describing the normalization, typing levels and logical 
mappings.
We also gave the patterns representing the most frequent properties, and 
commented on the effect of the coherence constraints on the conev
\end{document}